A Comparison of Parallel Graph Processing Implementations

### 2017-05-15 papers/publication/paper-pollard.tex
Updated some figures and reworked the tables.

\begin{table}
	\caption{A tabulated sample run of Graphalytics with 32 threads. Just one run per experiment is performed and these times do not take into account the different phases of execution. Runtimes are in seconds.}
	\centering
	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
	% dota-league (61 170 vertices, 50 870 313 edges)
	\begin{tabular}{l|c|c|c|c|c|c}
	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 1.0 & 11.1 & 13.0 &   5.0 & N/A  & 1.3 \\
	dota-league & 1.0 & 3.8 & 1104.0 & 2.5 & \textbf{2.9} & 0.9 \\
	\multicolumn{1}{l}{}\\
	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	\multicolumn{1}{l}{}\\
	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
	\end{tabular}

@@

\todo{redo this table}
\begin{table}
	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments. Performance results are in seconds running with the maximal number of threads (64 for PowerGraph, 72 for the others). Community detection is performed using label propagation.}
	\centering
	\pgfplotstabletypeset[
	col sep=comma,
	columns={[index]0,graphmat,openg,powergraph},
	every head row/.style={after row=\midrule},
	columns/0/.style={string type, column type={l|}, column name={}},
	columns/graphmat/.style={
		column name={GraphMat},
		string replace={0}{},
		dec sep align,
		empty cells with={N/A}
	},
	columns/openg/.style={
		column name={GraphBIG},
		dec sep align},
	columns/powergraph/.style={
		column name={PowerGraph},
		dec sep align}
	]{../report/runtime.csv}
	\label{tab:graphalytics}
\end{table}

@@

\subsection{Scalability}
\todo{People want more explanation of this}Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, a challenge facing most parallel graph algorithms. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
\begin{figure}
	\centering
	\begin{minipage}{\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup.pdf}
	\end{minipage}
	\begin{minipage}{\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphBIG dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}

\subsection{Real World Datasets}
\todo{Plot this}
\begin{table}
	\centering
	\caption{This table shows averages of runtimes for the Dota-League dataset using our method using 32 threads. Each cell represents the average over 32 runs. The ``N/A'' for PowerGraph and GraphBIG exist because both of these systems read in the files and build a graph data structure simultaneously.}
	\begin{tabular}{l|c|c|c|c}
		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
		\\ \hline
		GraphBIG   & N/A   & 0.0476  & 0.162    & 21.39  \\
		PowerGraph & N/A   & 0.36874 & 0.344    & 3.441  \\
		GAP        & 0.412 & 0.00464 & 0.267    & 0.2464 \\
		GraphMat   & 1.013 &         &          & 2.1587 \\
	\end{tabular}
	\label{tab:epg-perf}
\end{table}
%1         GAP      BFS Data structure build  0.25732000
%2    GraphMat      BFS Data structure build  2.10878359
%3         GAP PageRank Data structure build  0.32198688
%4    GraphMat PageRank Data structure build  2.15872441
%5         GAP     SSSP Data structure build  0.47479500
%6    GraphMat     SSSP Data structure build  2.35433178
%7         GAP      BFS                 Time  0.07449250
%8    GraphBIG      BFS                 Time  0.01132847
%9    GraphMat      BFS                 Time  0.02272806
%10        GAP PageRank                 Time  0.24646406
%11   GraphBIG PageRank                 Time 19.53251875
%12   GraphMat PageRank                 Time  1.69520531
%13 PowerGraph PageRank                 Time  3.44062500
%14        GAP     SSSP                 Time  0.07461594
%15   GraphBIG     SSSP                 Time  0.03791720
%16   GraphMat     SSSP                 Time  0.03936912
%17 PowerGraph     SSSP                 Time  0.36874997

>>>>

\begin{table}
	\caption{A tabulated sample run of Graphalytics with 32 threads. Just one run per experiment is performed and these times do not take into account the different phases of execution. Runtimes are in seconds.}
	\centering
	% cit-Patents: (3 774 768 vertices, 16 518 948 edges)
	% dota-league (61 170 vertices, 50 870 313 edges)
	\begin{tabular}{l|c|c|c|c|c|c}
	GraphBIG & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 1.0 & 11.1 & 13.0 &   5.0 & N/A  & 1.3 \\
	dota-league & 1.0 & 3.8 & 1104.0 & 2.5 & \textbf{2.9} & 0.9 \\
	\multicolumn{1}{l}{}\\
	PowerGraph & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 13.8 & 30.1 & 23.9 & 18.8 & N/A & 22.1 \\
	dota-league & 25.6 & 31.2 & 458.1 & 26.7 & 28.9 & 22.9 \\
	% graph500-22 43.0 s 55.6 s 299.8 s 46.4 s N/A    40.5 s
	\multicolumn{1}{l}{}\\
	GraphMat & BFS  & CDLP  &   LCC   &  PR &  SSSP & WCC \\ \hline
	cit-Patents & 7.8 & 20.6 & 11.2 &  8.4 & N/A &  7.0 \\
	dota-league & 2.2 & 19.7 & 150.1 & 2.8 & \textbf{3.8} & 3.1
	\end{tabular}

@@

\todo{redo this table}
\begin{table}
	\caption{The dataset used is the same Kronecker graph with scale $22$ as used in other experiments. Performance results are in seconds with 32 threads. Community detection is performed using label propagation.}
	\centering
	\pgfplotstabletypeset[
	col sep=comma,
	columns={[index]0,graphmat,openg,powergraph},
	every head row/.style={after row=\midrule},
	columns/0/.style={string type, column type={l|}, column name={}},
	columns/graphmat/.style={
		column name={GraphMat},
		string replace={0}{},
		dec sep align,
		empty cells with={N/A}
	},
	columns/openg/.style={
		column name={GraphBIG},
		dec sep align},
	columns/powergraph/.style={
		column name={PowerGraph},
		dec sep align}
	]{../report/runtime.csv}
	\label{tab:graphalytics}
\end{table}

@@

\subsection{Scalability}
\todo{People want more explanation of this}Figure~\ref{fig:bfs-scaling} shows the scalability and speedup for BFS. We define the strong scaling as $T_1 / (n T_n)$ where $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ is the time for $n$ threads. A linear scalability is when $T_n = T_1/n$ and is the horizontal line on the bottom plot in Fig.~\ref{fig:bfs-scaling}. These plots show in general poor scaling throughout, a challenge facing most parallel graph algorithms. Additionally, $2^{20}$ vertices is a ``small'' graph by today's standards and thus library designers focus on scalability for larger graphs. Another limitation may be the ability for OpenMP to efficiently handle such a large number of threads per machine. GCC version 4.8.5 supports OpenMP version 3.1 but the most recent release of OpenMP is 4.5.
\begin{figure}
	\centering
	\begin{minipage}{\linewidth}
		% trim=left lower right upper
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_speedup20.pdf}
	\end{minipage}
	\begin{minipage}{\linewidth}
		\includegraphics[width=\linewidth, trim=0 18pt 18pt 12pt, clip]{graphics/bfs_ss20.pdf}
	\end{minipage}
	\caption{The top figure has both axes logarithmic with an exception at 72 threads for readability. Linear speedup is. GraphBIG dips below 1 because it is slower for $2$ threads than for 1. $T_1$ is the serial time, $n$ is the number of threads, and $T_n$ time for $n$ threads.}
	\label{fig:bfs-scaling}
\end{figure}

\subsection{Real World Datasets}
\begin{table}
	\centering
	\caption{This table shows averages of runtimes for the Dota-League dataset using our method with 32 threads. We average over 32 runs and since the data structure construction times do not differ significantly we average across all algorithms. Both PowerGraph and GraphBIG read in the files and build a graph data structure simultaneously so we cannot measure data structure construction time.}
	\begin{tabular}{l|c|c|c|c}
		System & \begin{tabular}[x]{@{}c@{}}Construct Data \\ Structure\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ BFS\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ SSSP\end{tabular}
		& \begin{tabular}[x]{@{}c@{}}Run \\ PageRank\end{tabular}
		\\ \hline
		GraphBIG   & N/A   & 0.01133        & {\bf 0.03791} & 19.5325  \\
		PowerGraph & N/A   & N/A            & 0.3687        & 3.4406 \\
		GAP        & 0.351 & 0.07449        & 0.0746        & {\bf 0.2465} \\
		GraphMat   & 2.207 & {\bf 0.02273 } & 0.0397        & 1.6952 \\
	\end{tabular}
	\label{tab:epg-perf}
\end{table}
%1         GAP      BFS Data structure build  0.25732000
%2    GraphMat      BFS Data structure build  2.10878359
%3         GAP PageRank Data structure build  0.32198688
%4    GraphMat PageRank Data structure build  2.15872441
%5         GAP     SSSP Data structure build  0.47479500
%6    GraphMat     SSSP Data structure build  2.35433178
%7         GAP      BFS                 Time  0.07449250
%8    GraphBIG      BFS                 Time  0.01132847
%9    GraphMat      BFS                 Time  0.02272806
%10        GAP PageRank                 Time  0.24646406
%11   GraphBIG PageRank                 Time 19.53251875
%12   GraphMat PageRank                 Time  1.69520531
%13 PowerGraph PageRank                 Time  3.44062500
%14        GAP     SSSP                 Time  0.07461594
%15   GraphBIG     SSSP                 Time  0.03791720
%16   GraphMat     SSSP                 Time  0.03936912
%17 PowerGraph     SSSP                 Time  0.36874997
